\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}

\name{Jayden Thadani}
\class{MATH 139 --- Fourier Analysis}
\assignment{Homework \#13}
\duedate{9th May 2025}

\begin{document}

\begin{problem}[S \& S \S 8.4 \#7(b)]
	Recall that for $\lvert z \rvert < 1$,
	\[
		\log_{1} \left ( \frac{1}{1 - z} \right ) = \sum_{k \ge 1} \frac{z^{k}}{k}.
	\]
	We have seen that
	\[
		e^{\log_{1} (1 / (1 - z))} = \frac{1}{1 - z}.
	\]
	Show that if $\operatorname{Re} w > 1 /2$ and $w = \rho e^{i \varphi}$ with $\rho > 0$, $\lvert \varphi \rvert < \pi$, then
	\[
		\log_{1} w = \log \rho + i \varphi.
	\]
\end{problem}

\begin{solution}
	By part (a) of the same problem, since $\operatorname{Re} w > 1 / 2$, we can write
	\[
		w = \frac{1}{1 - z}
	\]
	with $\lvert z \rvert < 1$. Thus, we have $e^{\log_{1} w} = w$.

	If we write $\log_{1} w = a + b i$ and write $w$ in the polar coordinates given, we get $e^{a + b i} = e^{\log \rho + i \varphi}$. Comparing magnitudes, we can ignore the imaginary exponents to get $e^{a} = e^{\log \rho}$. Since $e^{x}$ is injective on $\mathbb{R}$, we have $a = \log \rho$.

	This forces $e^{bi} = e^{i \varphi}$, and thus, $bi - i \varphi = 2 \pi i m$ where $m$ is an integer-valued continuous function of $w$ since $\log_{1}$ is a continuous function. Since it takes values in $\mathbb{Z}$, it must in fact be constant. Thus, it suffices to show $m = 0$ for any one $w$. Consider $w = 0$ and write $w = 1 / (1 - 0)$. The polar form we have is $w = e^{0}$, and by the formula given, $\log_{1} w = \sum_{k} 0 = 0$. That is, the polar form and $\log_{1}$ coincide for $w = 0$ --- we have $m = 0$.
\end{solution}

\pagebreak

\begin{problem}[S \& S \S 8.4 \#11]
	Use the characters for $\mathbb{Z} / 3 \mathbb{Z}^{*}$, $\mathbb{Z} / 4 \mathbb{Z}^{*}$, $\mathbb{Z} / 5 \mathbb{Z}^{*}$, and $\mathbb{Z} / 6 \mathbb{Z}^{*}$ to verify directly that $L(1, \chi) \neq 0$ for all non-trivial Dirichlet characters modulo $q$ when $q = 3, 4, 5, 6$.
\end{problem}

\begin{solution}
	Recall (from \S 8.4 \# 6, on the previous homework) that $\phi(3) = \phi(4) = \phi(6)$ so we have \[
		\mathbb{Z} / 3 \mathbb{Z}^{*} \cong \mathbb{Z} / 4 \mathbb{Z}^{*} \cong \mathbb{Z} / 6 \mathbb{Z}^{*} \cong \mathbb{Z} / 2 \mathbb{Z}.
	\]
	In fact all of them consist of $1$ and its additive inverse in the corresponding addition groups. Thus, they have exactly one non-trivial character, defined by $\chi(-1) = -1$.

	For characters on $\mathbb{Z} / 3 \mathbb{Z}^{*}$ we get
	\[
		L(1, \chi) = 1 - \frac{1}{2} + \frac{1}{4} - \frac{1}{5} + \dots = \sum_{n \equiv 1}^{\infty} \frac{1}{n} - \frac{1}{n + 1}
	\]
	which is a sum of positive terms, and thus, is non-zero. Similarly on $\mathbb{Z} / 4 \mathbb{Z}^{*}$ we get
	\[
		L(1, \chi) = 1 - \frac{1}{3} + \frac{1}{5} - \frac{1}{7} + \dots = \sum_{n \equiv 1}^{\infty} \frac{1}{n} - \frac{1}{n + 2}
	\]
	and in $\mathbb{Z} / 6 \mathbb{Z}^{*}$ 
	\[
		L(1, \chi) = 1 - \frac{1}{5} + \frac{1}{7} - \frac{1}{11} + \dots = \sum_{n \equiv 1}^{\infty} \frac{1}{n} - \frac{1}{n + 4}
	\]
	both of which are sums of positive terms, and thus, positive.

	Since $5$ is a prime, we have $\mathbb{Z} / 5 \mathbb{Z}^{*} \cong \mathbb{Z} / 4 \mathbb{Z}$. In particular, $\mathbb{Z} / 5 \mathbb{Z}^{*} = \{ \overline{1}, \overline{2}, \overline{3}, \overline{4} \}$ is generated by $\overline{2}$. Thus, a character on $\mathbb{Z} / 5 \mathbb{Z}^{*}$ is completely determined by its value on $\overline{2}$. Since $\overline{2}$ is an order $4$ element, it can be mapped to elements of order $2$ or order $4$ in $S^{1}$. This gives $2$ characters --- defined by $\chi_{2}(\overline{2}) = -1$ and $\chi_{4}(\overline{2}) = i$.

	This gives us
	\[
		L(1, \chi_{2}) = 1 - \frac{1}{2} - \frac{1}{3} + \frac{1}{4} + \frac{1}{6} - \frac{1}{7} - \frac{1}{8} + \frac{1}{9} + \dots = \sum_{n \equiv 1} \left ( \frac{1}{n} - \frac{1}{n + 1} \right ) - \left ( \frac{1}{n + 3} - \frac{1}{n + 2} \right ).
	\]
	Since the difference between $1 / n$ and $1 / (n + 1)$ is greater than the difference between $1 / (n + 2)$ and $1 / (n + 3)$, this is a sum of positive terms, and thus, positive.

	We also have
	\[
		L(1, \chi_{4}) = 1 + \frac{i}{2} - \frac{1}{3} - \frac{i}{4} + \frac{1}{6} + \frac{i}{7} - \frac{1}{8} - \frac{i}{9} + \dots = \sum_{n \equiv 1} \left ( \frac{1}{n} - \frac{1}{n + 2} \right ) + \left ( \frac{1}{n + 1} - \frac{1}{n + 3} \right ) i
	\]
	which is the sum of positive real and imaginary parts, and thus, positive.
\end{solution}

\pagebreak

\begin{problem}[S \& S \S 8.4 \#12]
	Suppose $\chi$ is real and non-trivial; assuming the theorem that $L(1, \chi) \neq 0$, show directly that $L(1, \chi) > 0$.
\end{problem}

\begin{solution}
	If $\chi$ is real and non-trivial, then $\chi(p) = \pm 1$ for all $p \in \wp$. Since we're given that $L(1, \chi) \neq 0$, it suffices to show that $L(1, \chi) \ge 0$. Since $L(s, \chi)$ is continuous (and in fact $\mathcal{C}^{1}$) in $s$, we can do so by showing that $L(s, \chi) \ge 0$ even as $s \to 1$.

	The Dirichlet product formula gives
	\[
		L(s, \chi) = \prod_{p \in \wp} \frac{1}{1 - \chi(p) / p^{s}}
	\]
	Since $\chi(p) = \pm 1$, the denominator $1 - \chi (p) / p^{s}$ is positive. Since $L(s, \chi)$ is the product of positive terms, it must be non-negative. That is, $L(s, \chi) \ge 0$ for all $s > 1$. Taking the limit as $s \to 1$ it follows that $L(1, \chi) \ge 0$. Then since $L(1, \chi) \neq 0$, we have $L(1, \chi) > 0$ as desired.
\end{solution}

\end{document}
