\documentclass[12pt, letterpaper]{hmcpset}
\usepackage{lin-alg-preamble}

\name{Jayden Thadani}
\class{Linear Algebra Done Right self-study}
\assignment{Exercises 6.B --- Orthonormal Bases}

\begin{document}

\begin{problem}[1]
	\begin{enumerate}
		\item
			Suppose $\theta \in \bb{R}$. Show that $(\cos{\theta}, \sin{\theta}),$ $(-\sin{\theta}, \cos{\theta})$ and $(\cos{\theta}, \sin{\theta}),$ $(\sin{\theta}, - \cos{\theta})$ are orthonormal bases of $\bb{R}^2$.
		\item
			Show that each orthonormal basis of $\bb{R}^2$ is of the form given by one of the two possibilities of part (a).
	\end{enumerate}
\end{problem}

\begin{solution}
	\begin{enumerate}
		\item
			We can show that both of the lists are orthonormal lists of the length $\dim \bb{R}^2 = 2$, and thus, bases.\\\\
			First, note that for $v = (\pm \cos{\theta}, \pm \sin{\theta})$
			\[
				v \cdot v = \cos^2{\theta} + \sin^2{\theta} = 1.
			\]
			Thus, for either of the lists $e_1, e_2$, we have that $e_j \cdot e_j = 1$ ($j \in \set{1, 2}$).\\\\
			Further, we can see that 
			\[
				(\cos{\theta}, \sin{\theta}) \cdot (- \sin{\theta}, \cos{\theta}) = \cos{\theta} \sin{\theta} - \cos{\theta} \sin{\theta} = 0
			\]
			and
			\[
				(\cos{\theta}, \sin{\theta}) \cdot (\sin{\theta}, -\cos{\theta}) = -(\cos{\theta}, \sin{\theta}) \cdot (- \sin{\theta}, \cos{\theta}) = 0.
			\]
			That is, for either of the lists, we have $e_1 \cdot e_2 = 0$.\\\\
			Thus, both of the lists are orthonormal. Since both of the lists are of length $2 = \dim \bb{R}^2$, they are also orthonormal bases.
		\item
			Suppose we have an orthonormal basis of $\bb{R}^2$, say, $e_1, e_2$.\\\\
			Then $e_1$ must be a unit vector in $\bb{R}^2$. For angle $\theta$ from the positive $x$-axis, we can write $e_1$ as $(\cos{\theta}, \sin{\theta})$. Since we are working with the Euclidean inner product, $e_2$ must then be $e_1$ rotated $\frac{\pi}{2}$ clockwise or counterclockwise:
			\[
				e_2 = (\cos{\theta + \frac{\pi}{2}}, \sin{\theta + \frac{\pi}{2}}) = (-\sin{\theta}, \cos{\theta})
			\]
			or
			\[
				e_2 = (\cos{\theta - \frac{\pi}{2}}, \sin{\theta - \frac{\pi}{2}}) = (\sin{\theta}, -\cos{\theta}).
			\]
			This gives us any arbitrary orthonomal basis of $\bb{R}^2$ in the desired form.
	\end{enumerate}
\end{solution}

\begin{problem}[2]
	Suppose $e_1, \ldots, e_m$ is an orthonormal list of vectors in $V$. Let $v \in V$. Prove that
	\[
		\norm{v}^2 = \abs{\langle v, e_1 \rangle}^2 + \cdots + \abs{\langle v, e_m \rangle}^2
	\]
	if and only if $v \in \vspan{(e_1, \ldots, e_m)}$.
\end{problem}

\begin{problem}[3]
	Suppose $T \in \curly{L}(\bb{R}^3)$ has an upper-triangular matrix with respect to the basis $(1, 0, 0), (1, 1, 1), (1, 1, 2)$. Find an orthonormal basis of $\bb{R}^3$ (use the usual inner product on $\bb{R}^3$) with respect to which $T$ has an upper-triangular matrix.
\end{problem}

\begin{problem}[4]
	Suppose $n$ is a positive integer. Prove that
	\[
		\frac{1}{\sqrt{2 \pi}}, \frac{\cos{x}}{\sqrt{\pi}}, \frac{\cos{2x}}{\sqrt{\pi}}, \cdots, \frac{\cos{nx}}{\sqrt{\pi}}, \frac{\sin{x}}{\sqrt{\pi}}, \frac{\sin{2x}}{\sqrt{\pi}}, \cdots, \frac{\sin{nx}}{\sqrt{\pi}}
	\]
	is an orthonormal list of vectors in $\curly{C}[-\pi, \pi]$, the vector space of continuous functions on $[-\pi, \pi]$ with inner product
	\[
		\linn f, g \rinn =  \int_{-\pi}^{\pi} f(x) g(x) \, dx.
	\]
\end{problem}

\begin{problem}[5]
	On $\curly{P}_2(\bb{R})$, consider the inner product given by
	\[
		\linn p, q \rinn \int_0^1 p(x) q(x) \, dx.
	\]
	Apply the Gram-Schmidt procedure to the basis $1, x, x^2$ to produce an orthonormal basis of $\curly{P}_2(\bb{R})$
\end{problem}

\begin{problem}[6]
	Find an othonormal bass of $\curly{P}_2(\bb{R})$ (with inner product as in exercise 5) such that the differentiation operator (the operator that takes $p$ to $p'$) has an upper-triangular matrix with respect to this basis.
\end{problem}

\begin{problem}[7]
	Find a polynomial $q \in \curly{P}_2(\bb{R})$ such that
	\[
		p(\frac{1}{2}) = \int_0^1 p(x) q(x) \, dx
	\]
	for every $p \in \curly{P}_2(\bb{R})$.
\end{problem}

\begin{problem}[8]
	Find a polynomial $q$ such that
	\[
		\int_0^1 p(x) \cos{\pi x} \, dx = \int_0^1 p(x) q(x) \, dx
	\]
	for every $p \in \curly{P}_2(\bb{R})$.
\end{problem}

\begin{problem}[9]
	What happens if the Gram-Schmidt procedure is applied to a list of vectors that is not linearly independent?
\end{problem}

\begin{problem}[10]
	Suppose $V$ is a real inner product space and $v_1 \ldots v_m$ is a linearly independent list of vectors in $V$. Prove that there exist exactly $2^m$ orthonormal lists $e_1, \ldots, e_m$ of vectors in $V$ such that
	\[
		\vspan (v_1, \ldots, v_j) = \vspan (e_1, \ldots, e_j)
	\]
	for all $j \in \bb{N}_m$
\end{problem}

\end{document}


