\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}

\name{Jayden Thadani}
\class{MATH 147 --- Topology}
\assignment{Reflection exercise}
\duedate{27th March 2025}

\begin{document}

\begin{problem}
	Reflect on the following question:

	What have you learned in this class about the process of doing or creating mathematics?

	Weave in potential answers to the following:
	\begin{itemize}
		\item have you experienced: joy, beauty, reward?
		\item which theorems have you proved that you feel particularly proud of?
		\item what have you learned about the value of struggle?
		\item what have you learned about the importance of community?
	\end{itemize}
\end{problem}

\begin{solution}
	I've enjoyed this class as a whole. I really like topology as a subject because of the beautiful results it produces and its set-theoretic nature. I've also been very happy with the structure of the class --- proving results on my own and working on presentation skills has been a very valuable experience for me.

	I think my biggest takeaway from this class is that I really can do math on my own. Often, in other classes where outside resources are less restricted, I use  resources like StackExchange to figure out if small lemmas I can't be bothered to prove exist, or to check counterexamples to methods that I am trying. The restricted resources of this course have forced me to rely on my own intuition for whether an approach will work or not.

	One recent example that I'm really proud of is the proof that continuous functions are determined on dense sets. Originally I came up with a sequential proof that seemed to work on codomain spaces where limit points of a set have a sequence of points converging to them (1st countable spaces). However, I was a lot less sure that this held for the more general class of all Hausdorff codomains that we were supposed to prove it for. Usually, I would have looked up examples of Hausdorff spaces and tried to see if this approach worked. Instead, I just tried different approaches to the problem for a week and a half, way after I had already submitted my first approach. Finally it hit me --- if $f, g : X \to Y$ are continuous functions from a general topological space to a Hausdorff topological space that agree on the dense subset $D \subseteq X$, I could show that $f(p) = g(p)$ for any $p \in X$ by showing that $f(p)$ and $g(p)$ cannot be separated with disjoint open neighbourhoods. Even though this result and this proof are not in themselves particularly impressive, I'm proud that I thought about the problem on my own for so long and eventually solved it.

	Finally I'm really grateful for the community we've formed in the topology class. It's been lovely to have a community that provides the feedback on my solutions and presentation, and with whom I can share my thoughts about topology. I can honestly say that I've made and strengthened friendships in this class.
\end{solution}

\end{document}
