\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}

\name{Jayden Thadani}
\class{MATH 189AA --- Commutative Algebra}
\assignment{Exam 2}
\duedate{25th November 2025}

\begin{document}

\begin{problem}[1.]
	Consider the ideal $I = (x, z)$ in the ring
	\[
		R = \mathbb{Q}[x, y, z] / (y^{2} - xz, x^{2} y - z^{2}, x^{3} - yz).
	\]
	\begin{enumerate}
		\item Use Macaulay2 to calculate a free resolution $P_{\bullet}$ of $I$ over $R$. Find the modules $P_{0}, P_{1}, P_{2}$ and the matrices representing the homomorphisms between them. Write your solution in the form
			\[
				\begin{tikzcd}
					\cdots \ar[r] & P_{2} \ar[r, "\text{matrix}"] & P_{1} \ar[r, "\text{matrix}"] & P_{0} \ar[r, "\text{matrix}"] & (x, z) \ar[r] & 0
				\end{tikzcd}
			\]

		\item Print your Macaulay2 session and hand it in with your exam.
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[2.]
	The diagram below is a commutative diagram of $R$-modules with exact rows:
	\[
		\begin{tikzcd}
			A \ar[r, "f_{1}"] \ar[d] & B \ar[r, "f_{2}"] \ar[d] & C \ar[r, "f_{3}"] \ar[r] \ar[d] & D \ar[r, "f_{4}"] \ar[d] & E \ar[d] \\
			A' \ar[r, "f_{1}"] & B' \ar[r, "f_{2}"] & C' \ar[r, "f_{3}"] \ar[r] & D' \ar[r, "f_{4}"] & E'
		\end{tikzcd}
	\]
	Establish the following claims directly, by diagram chase:
	\begin{enumerate}
		\item If $\alpha$ is surjective and $\beta$ and $\delta$ are both injective, then $\gamma$ is injective.

		\item If $\varepsilon$ is injective and $\beta$ and $\delta$ are both surjective, then $\gamma$ is surjective.
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[3.]
	If $P$ is a finitely generated projective $R$-module, show that $\operatorname{Hom}_{R}(P, R)$ is also a projective $R$-module.
\end{problem}

\begin{solution}
	Since $P$ is projective, we can find some $R$-module $N$ such that $P \oplus N \cong F$ is free. Thus,
	\[
		\operatorname{Hom}_{R}(F, R) \cong \operatorname{Hom}_{R}(P, R) \times \operatorname{Hom}_{R}(N, R) \cong \operatorname{Hom}_{R}(P, R) \oplus \operatorname{Hom}_{R}(N, R).
	\]
	Further, since $P$ is finitely generated, we can choose $F$ to be free. We claim that this makes $\operatorname{Hom}_{R}(F, R)$ free. Then, since we have written $\operatorname{Hom}_{R}(P, R)$ as a direct summand of a free module, we have shown that it is projective.

	The freeness of $\operatorname{Hom}_{R}(F, R)$ follows by the chain of isomorphisms below.
	\begin{align*}
		\operatorname{Hom}_{R}(F, R) & \cong \operatorname{Hom}_{R}(R^{\oplus n}, R) \\
					     & = \bigoplus_{i = 1}^{n} \operatorname{Hom}_{R}(R, R) \\
					     & = R^{\oplus n}.
	\end{align*}
	The last equality follows because $\operatorname{Hom}_{R}(R, M) \cong M$.
\end{solution}

\pagebreak

\begin{problem}[4.]
	Let
	\[
		\begin{tikzcd}
			0 \ar[r] & L \ar[r] & M \ar[r] & N \ar[r] & 0
		\end{tikzcd}
	\]
	be a short exact sequence of $R$-modules. If $M$ and $N$ are projective, prove that $L$ is also projective.
\end{problem}

\begin{solution}
	Since $N$ is projective, the exact sequence splits. Thus, we have $M \cong L \oplus N$. Since $M$ is projective, it is a direct summand of some free module $F$. But then, as a direct summand of $M$, $L$ also becomes a direct summand of $F$ (if $M \oplus Q \cong F$, then $L \oplus (N \oplus Q) \cong F$). That is, $L$ is projective.
\end{solution}

\pagebreak

\begin{problem}[5.]
	Suppose $M$ is a finitely presented $R$-module and let
	\[
		\begin{tikzcd}
			0 \ar[r] & K \ar[r] & N \ar[r] & M \ar[r] & 0
		\end{tikzcd}
	\]
	be a short exact sequence where $N$ is finitely generated. Prove that $K$ is also finitely generated.
\end{problem}

\begin{solution}
\end{solution}

\end{document}
