\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{lin-alg-preamble}

\name{Jayden Thadani}
\class{MATH 171 --- Abstract Algebra I}
\assignment{Homework 8}
\duedate{16th February 2024}

\begin{document}

\begin{problem}[Dummit \& Foote \S2.1 \#6]
	Let $G$ be an abelian group. Prove that $\{ g \in G | \lvert g \rvert < \infty \}$ is a subgroup of $G$ (called the \textit{torsion subgroup} of $G$). Give an example where this set is not a subgroup when $G$ is non-abelian.
\end{problem}

\begin{solution} 
	Suppose $G$ is an arbitrary abelian group. Let $H$ denote the set of all elements of finite order. We want to show that $H$ is a subgroup of $G$. Note that we only need to show that for all $x, y \in H$ we have $xy^{-1} \in G$ (and that $H$ is nonempty). \\\\
	We have trivially that $H$ is nonempty since $1_G$ has finite order. Suppose $x, y \in H$. Specifically, suppose $\lvert x \rvert = m$ and $\lvert y \rvert = n$ with $m, n \in \mathbb{N}$. Then note that
	\[
		\begin{split}
			(xy^{-1})^{mn} & = \underbrace{(xy^{-1}) \dots (xy^{-1})}_{mn \text{ times}} \\
				       & = x^{mn} (y^{-1})^{mn} \\
				       & = (x^{m})^n (y^n)^{-m} \\
				       & = 1_G.
		\end{split}
	\]
	Thus, $\lvert xy^{-1} \rvert \le mn$ and $xy^{-1} \in H$ as desired.
\end{solution}

\pagebreak

\begin{problem}[Dummit \& Foote \S2.1 \#8]
	Let $H$ and $K$ be subgroups of $G$. Prove that $H \cup K$ is a subgroup if and only if either $H \subset K$ or $K \subset H$.
\end{problem}

\pagebreak

\begin{problem}[Dummit \& Foote \S2.1 \#10]
	\begin{enumerate}
		\item Prove that if $H$ and $K$ are subgroups of $G$ then so is their intersection $H \cap K$.
		\item Prove that the intersection of an arbitrary nonempty collection of subgroups of $G$ is again a subgroup of $G$ (do not assume the collection is countable).
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[Dummit \& Foote \S2.1 \#15]
	Let $H_1 \le H_2 \le \dots$ be an ascending chain of subgroups of $G$. Prove that $\bigcup^\infty_{i = 1} H_i$ is a subgroup of $G$.
\end{problem}

\end{document}
