\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}
\usepackage{amsthm}

\name{Jayden Thadani}
\class{MATH 171 --- Abstract Algebra}
\assignment{Exam 2}
\duedate{2nd April 2024}

\newtheorem*{theorem}{Theorem}

\begin{document}

\begin{problem}[1.]
	Let $A$ and $B$ be subsets of $S_{4}$ where $A = \{ (1 \, 2), (3 \, 4) \}$ and $B = \{ (1 \, 2)(3 \, 4) \}$. Prove or disprove that $N_{S_{4}}(A)$ is a subgroup of $N_{S_{4}}(B)$.
\end{problem}

\pagebreak

\begin{problem}[2.]
	Prove that $S_{4}$ does not have a normal subgroup of order $8$.
\end{problem}

\pagebreak

\begin{problem}[3.]
	Write up a proof of the 3rd isomorphism theorem.
\end{problem}
	\begin{theorem}
		For any normal subgroups $H, K$ of $G$, with $H \le K$, we have
		\[
			K/H \trianglelefteq G/H
		\]
		and
		\[
			(G/H)/(K/H) \cong G/H. 
		\]
	\end{theorem}
	\begin{proof}
		We claim that the map
		\[
			\begin{split}
				\varphi & : G/H \to G/K \\
					  & : gH \mapsto gK
			\end{split}
		\]
		is a surjective homomorphism with kernel $K/H$ (and that thus, by the first isomorphism theorem we have $K/H \trianglelefteq G/H$ and $(G/H)/(K/H) \cong G/K$). \\\\
		First we show that $\varphi$ is well-defined. Suppose $g_{1} H = g_{2} H$. Then $g_{1} g_{2}^{-1} \in H$. Since $H \le K$, this gives us that $g_{1} g_{2}^{-1} \in K$, and thus $g_{1} K = g_{2} K$. That is, $\varphi(g_{1} H) = \varphi(g_{2} H)$, and thus, $\varphi$ is well-defined. \\\\
		Now we show that $\varphi$ is a homomorphism. To do so we can just compute $\varphi((g_{1}H)(g_{2} H))$ for any $g_{1}H, g_{2}H \in G/H$.
		\[
			\begin{split}
				\varphi((g_{1}H)(g_{2}H)) & = \varphi((g_{1}g_{2})H) \\
							  & = (g_1 g_{2}) K \\
							  & = (g_{1} K) (g_{2} K) \\
							  & = \varphi(g_{1} H) \varphi(g_{2} H).
			\end{split}
		\]
		This proves that $\varphi$ is a homomorphism. \\\\
		For any $gK \in G/K$, we know that there exists $gH \in G/H$, and by definition, $\varphi(gH) = gK$. Thus, $\varphi$ contains every $gK \in G/K$ in its image. That is, $\varphi(G) = G/K$. \\\\
		Finally, we will show that $\ker \varphi = K/H$ by double containment. \\\\
		Suppose $gH \in \ker \varphi$. Then $\varphi(gH) = 1_{G/K} = K$. But we know $\varphi(gH) = gK$. Thus, we must have $gK = K$, and thus $g \in K$. Then $gH \in K/H$. Thus $\ker \varphi \subset K/H$.  \\\\
		Suppose $kH \in K/H$ where $k \in K$. Then
		\[
			\begin{split}
				\varphi(kH) & = kK \\
					    & = kK \\
					    & = K \\
					    & = 1_{K/H}.
			\end{split}
		\]
		Thus, $kH \in \ker \varphi$. Thus, $K/H \subset \ker \varphi$. Then, by double containment. $K/H = \ker \varphi$. \\\\
		Finally, by the first isomorphism theorem $K/H \trianglelefteq G/H$ and $(G/H)/(K/H) \cong G/K$.
	\end{proof}

\pagebreak

\begin{problem}[4.]
	Let $H$ be a normal subgroup of a finite group $G$. Prove that if $G/H$ has an element of order $n$, then $G$ has an element of order $n$.
\end{problem}

\begin{solution}
	Let $\overline{g} = gH \in G/H$ be an element of order $n$. Thus, $\overline{g}^{j} \neq \overline{g}^{k}$ for any $j \neq k$ with $j, k \in \{ 1, \dots, n \}$ and $\overline{g}^{n} = H$. Thus, $g^{j} g^{-k} \notin H$ for all those $j, k$ and $g^{n} = h \in H$. \\\\
	Since $G$ is finite, we must have $g$ must also have finite order --- say $\lvert g \rvert = m$. Note that we must have $n \le m$ since if $m < n$ we'd have $g^{m + 1} g^{-1} = 1 \in H$ with $1, m + 1 \in \{ 1, \dots, n \}$. 
\end{solution}

\pagebreak

\begin{problem}[5.]
	Briefly explain why the following statements are true or false
	\begin{enumerate}
		\item If $K \trianglelefteq H$ and $H \trianglelefteq G$ then $K \trianglelefteq G$.
		\item If $H \le G$, then the left cosets of $H$ partition $G$.
		\item If $H \trianglelefteq G$, then $H$ is the kernel of some homomorphism with domain $G$.
		\item If $\lvert G \rvert < \infty$ and $n$ divides $\lvert G \rvert$, then $G$ has a subgroup $H$ such that $\lvert H \rvert = n$.
		\item If $H$ and $K$ are subgroups of $G$ then $HK$ is a subgroup of $G$.
	\end{enumerate}
\end{problem}

\begin{solution}
	\begin{enumerate}
		\item
		\item
		\item \fbox{True} \\\\
			Since $H \trianglelefteq G$, $G/H$ is a group with the standard binary operation. Consider the projection $\pi : G \to G/H$ by $g \mapsto gH$. Notice that $gH = 1_{G/H}$ if and only if $g \in H$. Thus $\ker \pi = H$.
	\end{enumerate}
\end{solution}

\end{document}
