\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}

\name{Jayden Thadani}
\class{MATH 181 --- Dynamical Systems}
\assignment{Homework 1}
\duedate{1st February 2026}

\begin{document}

\begin{problem}[Strogatz 2.2.7]
	Analyse the following equation graphically. Sketch the vector field on the real line, find all the fixed points, classify their stability, and sketch the graph of $x(t)$ for different initial conditions. Then try for a few minutes to obtain the analytical solution for $x(t)$.
	\[
		\dot{x} = e^{x} - \cos{x}.
	\]
\end{problem}

\pagebreak

\begin{problem}[Strogatz 2.2.8]
	Given an equation $\dot{x} = f(x)$, we know how to sketch the corresponding flow on the real line. Here you are asked to solve the opposite problem: for the phase portrait shown, find an equation that is consistent with it.
\end{problem}

\pagebreak

\begin{problem}[Strogatz 2.2.10]
	For each of (a)--(e), find an equation $\dot{x} = f(x)$ with the stated properties, or if there are no examples, explain why not. In all cases, assume that $f(x)$ is a smooth function.
	\begin{enumerate}
		\item Every real number is a fixed point.
		\item Every integer is a fixed point and there are no others.
		\item There are precisely three fixed points and all of them are stable.
		\item There are no fixed points.
		\item There are precisely $100$ fixed points.
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[Strogatz 2.3.4]
	For certain species of organisms, the effective growth rate $\dot{N} / N$ is highest at intermediate $N$. This is called the Allee effect. For example, imagine that it is too hard to find mates when $N$ is very small, and there is too much competition for food and other resources when $N$ is large.
	\begin{enumerate}
		\item Show that $\dot{N} / N = r - a(N - b)^{2}$ provides an example of the Allee effect, if $r$, $a$, and $b$ satisfy certain constraints, to be determined.

		\item Find all fixed points of the system and classify their stability.

		\item Sketch the solutions $N(t)$ for different initial conditions.

		\item Compare the solutions $N(t)$ for those found for the logistic equation. What are the qualitative differences, if any?
	\end{enumerate}
\end{problem}

\pagebreak

Use linear stability analysis to classify the fixed points of the following system. If linear stability analysis fails because $f'(x_{*}) = 0$, use a graphical argument to decide the stability.

\begin{problem}[Strogatz 2.4.4]
	\[
		\dot{x} = x^{2}(6 - x)
	\]
\end{problem}

\begin{problem}[Strogatz 2.4.7]
\[
	\dot{x} = ax - x^{3}
\]
where $a$ can be positive, negative, or zero. Discuss all three cases.
\end{problem}

\pagebreak

\begin{problem}[Strogatz 2.5.2]
	Show that the solution to $\dot{x} = 1 + x^{10}$ escapes to $+\infty$ in a finite time, starting from any initial condition.
\end{problem}

\pagebreak

\begin{problem}[Strogatz 2.6.2]
	Here's an analytic proof sketch that periodic solutions are impossible for a vector field on a line. Suppose, on the contrary, that $x(t)$ is a non-trivial periodic solution, for some $T > 0$, and $x(t) \neq x(t + s)$ for all $0 < s < T$. Derive a contradiction by considering
	\[
		\int_{t}^{t + T} f(x) \frac{d x}{d t} \, dt.
	\]
\end{problem}

\end{document}
