\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}

\name{Jayden Thadani}
\class{Basic category theory}
\assignment{Exercises 1.2 --- Functors}
\duedate{}

\begin{document}

\begin{problem}[1.2.20]
	Find three examples of functors not mentioned above.
\end{problem}

\begin{solution}
	\begin{enumerate}[label = (\arabic*)]
		\item The pushforward that sends manifolds $M$ to their tangent bundles $\mathrm{T} M$, and sends smooth maps $f : M \to N$ to their tangent maps $f_{*} : \mathrm{T} M \to \mathrm{T} N$ is a functor from the category of smooth manifolds to the category of tangent bundles.

			In general, pushforwards are (covariant) functors. For example, a continuous map $\pi : X \to Y$ sends a sheaf $\mathscr{F}$ on $X$ to a pushforward sheaf $\pi_{*} \mathscr{F}$ on $Y$ defined by $\pi_{*} \mathscr{F}(U) = \mathscr{F}(\pi^{\text{pre}}(U))$. A morphism of sheaves $\phi : \mathscr{F} \to \mathscr{G}$ on $X$ is mapped to a morphism of sheaves $\pi_{*} \phi : \pi_{*} \mathscr{O} \to \pi_{*} \mathscr{G}$ where if $\phi(U) : \mathscr{F}(U) \mapsto \mathscr{G}(U)$, then $\pi_{*} \phi : \pi_{*} \mathscr{F}(U) \mapsto \pi_{*} \mathscr{G}(U)$.
	\end{enumerate}
\end{solution}

\pagebreak

\begin{problem}[1.2.21]
	Show that functors preserve isomorphism. That is, prove that if $F : \mathscr{A} \to \mathscr{B}$ is a functor, and $A, A' \in \mathscr{A}$ with $A \cong A'$, then $F(A) \cong F(A')$.
\end{problem}

\begin{solution}
	If $A \cong A'$, then there exist morphisms $f : A \to A'$ and $f' : A' \to A$ with $f \circ f' = \operatorname{id}_{A}$ and $f' \circ f = \operatorname{id}_{A'}$. We claim $F(f)$ and $F(f')$ serve the same purpose for $F(A)$ and $F(A')$.

	In particular, $F(f) \circ F(f') : F(A) \to F(A)$ is just
	\[
		F(f \circ f') = F(\operatorname{id}_{A}) = \operatorname{id}_{F(A)}.
	\]
	Similarly, $F(f') \circ F(f) : F(A') \to F(A')$ is just 
	\[
		F(f' \circ f) = F(\operatorname{id}_{A'}) = \operatorname{id}_{F(A')}.
	\]
	Thus, $F(f) : F(A) \to F(A')$ is the desired isomorphism.
\end{solution}

\pagebreak

\begin{problem}[1.2.22]
	Prove the assertion made in Example 1.2.9. In other words, given ordered sets $A$ and $B$, and denoting by $\mathscr{A}$ and $\mathscr{B}$ the corresponding categories, show that a functor $\mathscr{A} \to \mathscr{B}$ amounts to an order-preserving map $A \to B$.
\end{problem}

\begin{solution}
	In a pre-order category $\mathscr{P}$ corresponding to the pre-ordered set $P$, there exists a morphism $f : p_{1} \to p_{2}$ if and only if $p_{1} \le p_{2}$ in $P$. A functor $F : \mathscr{A} \to \mathscr{B}$ has a map $F(a_{1}) \to F(a_{2})$ if and only if there is a map $a_{1} \to a_{2}$. Equivalently, on the corresponding pre-ordered sets, $F(a_{1}) \le F(a_{2})$ in $B$ if and only if $a_{1} \le a_{2}$ in $A$ --- $F$ is an order preserving map.
\end{solution}

\pagebreak

\begin{problem}[1.2.23]
	Two categories $\mathscr{A}$ and $\mathscr{B}$ are \emph{isomorphic}, written as $\mathscr{A} \cong \mathscr{B}$, if they are isomorphic as objects of $\mathsf{CAT}$.
	\begin{enumerate}
		\item Let $G$ be a group, regarded as a one-object category all of whose maps are isomorphisms. Then its opposite $G^{\text{op}}$ is also a one-object category all of whose maps are isomorphisms, and can therefore be regarded as a group too. What is $G^{\text{op}}$, in purely group-theoretic terms? Prove that $G$ is isomorphic to $G^{\text{op}}$.
		\item Find a monoid not isomorphic to its opposite.
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[1.2.24]
	Is there a functor $Z : \mathsf{Grp} \to \mathsf{Grp}$ with the property that $Z(G)$ is the centre of $G$ for all groups $G$?
\end{problem}

\pagebreak

\begin{problem}[1.2.25]
	Sometimes we meet functors whose domain is a product $\mathscr{A} \times \mathscr{B}$ of categories. Here you will show that such a functor can be regarded as an interlocking pair of families of functors, one defined on $\mathscr{A}$ and the other defined on $\mathscr{B}$. (This is very like the situation for bilinear and linear maps).
	\begin{enumerate}
		\item Let $F : \mathscr{A} \times \mathscr{B} \to \mathscr{C}$ be a functor. Prove that for each $A \in \mathscr{A}$, there is a functor $F^{A} : \mathscr{B} \to \mathscr{C}$ defined on objects $B \in \mathscr{B}$ by $F^{A}(B) = F(A, B)$ and on maps $g \in \mathscr{B}$ by $F^{A}(g) = F(\operatorname{id}_{A}, g)$. Prove that for each $B \in \mathscr{B}$, there is a functor $F_{B} : \mathscr{A} \to \mathscr{C}$ defined similarly.
		\item Let $F : \mathscr{A} \times \mathscr{B} \to \mathscr{C}$ be a functor. With notation as in (a), show that the families of functors $(F^{A})_{A \in \mathscr{A}}$ and $(F_{B})_{B \in \mathscr{B}}$ satisfy the following two conditions:
			\begin{enumerate}[label = (\roman*)]
				\item if $A \in \mathscr{A}$ and $B \in \mathscr{B}$, then $F^{A}(B) = F_{B}(A)$ 
				\item $f : A \to A'$ in $\mathscr{A}$ and $g : B \to B'$ in $\mathscr{B}$, then $F^{A'}(g) \circ F_{B}(f) = F_{B'}(f) \circ F^{A}(g)$.
			\end{enumerate}
		\item Now take categories $\mathscr{A}$, $\mathscr{B}$, and $\mathscr{C}$, and take families of functors $(F^{A})_{A \in \mathscr{A}}$ and $(F_{B})_{B \in \mathscr{B}}$ satisfying the two conditions in (b). Prove that there is a unique functor $F : \mathscr{A} \times \mathscr{B} \to \mathscr{C}$ satisfying the equations in (a).
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[1.2.26]
	Fill in the details of Example 1.2.11, thus constructing a functor  $C : \mathsf{Top}^{\text{op}} \to \mathsf{Ring}$.
\end{problem}

\begin{solution}
	We define $C$ on objects by defining $C(X)$ to be the ring of continuous real functions on $X$. 

	Consider any continuous map $f : X \to Y$. Then, each map $\varphi : Y \to \mathbb{R}$ induces a map $\varphi \circ f : X \to \mathbb{R}$. We claim that the pullback $f^{*} : C(Y) \to C(X)$ by $\varphi \mapsto \varphi \circ f$ is a ring homomorphism. This follows since for any $\varphi, \psi \in C(Y)$, we have
	\begin{align*}
		f^{*}(\varphi + \psi)(p) & = ((\varphi + \psi) \circ f)(p) \\
				      & = (\varphi + \psi)(f(p)) \\
				      & = \varphi(f(p)) + \psi(f(p)) \\
				      & = (\varphi \circ f)(p) + (\psi \circ f)(p) \\
				      & = f^{*} \varphi(p) + f^{*} \psi(p).
	\end{align*}
	at every point, giving $f^{*}(\varphi + \psi) = f^{*} \varphi + f^{*} \psi$. Similarly $f^{*} (\varphi \psi) = (f^{*} \varphi) (f^{*} \psi)$. Thus, we define $C(f) = f^{*}$. To show $C$ is a contravariant functor we need to show that $C$ commutes with composition (in the opposite direction) and preserves the identity.

	First, the identity $\operatorname{id}_{X}$ on any topological space $X$ has $C(\operatorname{id}_{X}) : \varphi \mapsto \varphi \circ \operatorname{id}_{X} = \varphi$, so $C(\operatorname{id}_{X}) = \operatorname{id}_{C(X)}$ as desired. For any continuous maps $f : X \to Y$ and $g : Y \to Z$, the continuous map $g \circ f$ has $C(g \circ f) : \varphi \mapsto \varphi \circ (g \circ f)$. By the associativity of composition we have
	\begin{align*}
		C(g \circ f)(\varphi) & = \varphi \circ (g \circ f) \\
				   & = (\varphi \circ g) \circ f \\
				   & = C(g)(\varphi) \circ f \\
				   & = C(f)(C(g)(\varphi)) \\
				   & = (C(f) \circ C(g))(\varphi).
	\end{align*}
	That is $C(g \circ f) = C(f) \circ C(g)$, so $C$ is a functor.
\end{solution}

\pagebreak

\begin{problem}
	Find an example of a functor $F : \mathscr{A} \to \mathscr{B}$ such that $F$ is faithful, but there exist distinct maps $f_{1}$ and $f_{2}$ in $\mathscr{A}$ with $F(f_{1}) = F(f_{2})$. 
\end{problem}

\pagebreak

\begin{problem}[1.2.28]
	\begin{enumerate}
		\item Of the examples of functors appearing in this section, which are faithful and which are full?
		\item Write down one example of a functor that is both full and faithful, one that is full but not faithful, one that is faithful but not full, and one that is neither.
	\end{enumerate}
\end{problem}

\pagebreak

\begin{problem}[1.2.29]
	\begin{enumerate}
		\item What are the subcategories of an ordered set? Which are full?
		\item What are the subcategories of a group? Which are full?
	\end{enumerate}
\end{problem}

\end{document}
