\documentclass[12pt,letterpaper]{hmcpset}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{graphicx} % Required for inserting images
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\DeclareMathOperator{\vspan}{span}
\newcommand{\bb}[1]{\mathbb{#1}}
\newcommand{\bvec}[1]{\mathbf{#1}}
\newcommand{\pardx}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ddx}[2]{\frac{d #1}{d #2}}
\newcommand{\norm}[1]{\Vert #1 \Vert}
\newcommand{\abs}[1]{\vert #1 \vert}
\newcommand{\set}[1]{\{ #1 \}}
\setlist[enumerate, 1]{label = (\alph*)}

\name{Jayden Thadani}
\class{Principles of Mathematical Analysis self-study}
\assignment{Chapter 2 exercises---Basic Topology}

\begin{document}

\begin{problem}[1.]
	Prove that the empty set is a subset of every set.
\end{problem}

\begin{problem}[2.]
	A complex number $z$ is said to be \textit{algebraic} if there are integers $a_0, \ldots, a_n$, not all zero, such that
	\[
		a_0 z^n + a_1 z^{n - 1} + \cdots + a_{n - 1}z + a_n = 0.
	\]
	Prove that the set of algebraic numbers is countable. \textit{Hint:} For every positive integer $N$ there are only finitely many equations with
	\[
		n + \abs{a_0} + \abs{a_1} + \cdots + \abs{a_n} = N.
	\]
\end{problem}

\begin{problem}[3.]
	Prove that there exist real numbers which are not algebraic.
\end{problem}

\begin{problem}[4.]
	Is the set of all irrational real numbers countable?
\end{problem}

\begin{problem}[5.]
	Construct a bounded set of real numbers with exactly three limit points.
\end{problem}

\begin{problem}[6.]
	Let $E'$ be the set of all limit points of a set $E$. Prove that $E'$ is closed. Prove that $E$ and $\bar{E}$ have the same limit points. (Recall that $\bar{E} = E \cup E'$). Do $E$ and $E'$ always have the same limit points.
\end{problem}

\begin{problem}[7.]
	Let $A_1, A_2, A_3, \ldots$ be subsets of a metric space.
	\begin{enumerate}
		\item If $B_n = \bigcup_{i = 1}^n A_i$ prove that $\bar{B_n} = \bigcup_{i = 1}^n \bar{A_i}$.
	\item If $B_n = \bigcup_{i = 1}^\infty$ prove that $\bar{B_n} \supset \bigcup_{i = 1}^\infty \bar{A_i}$
	\end{enumerate}
\end{problem}

\begin{problem}[8.]
	Is every point of every open set $E \subset \bb{R}^2$ a limit point of $E$? Answer the same question for closed sets in $\bb{R}^2$.
\end{problem}

\begin{problem}[9.]
	Let $E^\circ$ denote the set of all interior points of a set $E$. $E^\circ$ is called the \textit{interior} of $E$.
	\begin{enumerate}
		\item Prove that $E^\circ$ is always open.
		\item Prove that $E$ is open if and only if $E^\circ = E$.
		\item If $G \subset E$ and $G$ is open, prove that $G \subset E^\circ$
		\item Prove that the complement of $E^\circ$ is the closure of the complement of $E$.
		\item Do $E$ and $\bar{E}$ always have the same interiors?
		\item Do $E$ and $E^\circ$ always have the same closures?
	\end{enumerate}
\end{problem}

\begin{problem}[10.]
	Let $X$ be an infinite set. For $p \in X$ and $q \in X$, define
	\[
		d(p, q) = \begin{cases}
			1 & \text{( if $p \neq q$)} \\
			0 & \text{( if $p = q$)}
		\end{cases}
	\]
\end{problem}

\begin{problem}[11.]
	For $x, y \in \bb{R}$, define
	\[
		\begin{gathered}
			d_1(x, y) = (x - y)^2 \\
			d_2(x, y) = \sqrt{\abs{x - y}} \\
			d_3(x, y) = \abs{x^2 - y^2} \\
			d_4(x, y) = \abs{x = 2y} \\
			d_5(x, y) = \frac{\abs{x - y}}{1 + \abs{x - y}}.
		\end{gathered}
	\]
	Determine, for each of these, whether it is a metric or not.
\end{problem}

\begin{problem}[12.]
	Let $K \subset \bb{R}$ consist of $0$ and the real numbers $1/n$ for $n \in \bb{N}$. Prove that $K$ is compact directly from the definition (without using the Heine-Borel theorem).
\end{problem}

\begin{problem}[13.]
	Construct a compact set of real numbers whose limit points form a countable set.
\end{problem}

\begin{problem}[14.]
	Give an example of an open cover of the segement $(0, 1)$ which has no finite subcover.
\end{problem}

\begin{problem}[15.]
	Show that theorem 2.36 and its corollary become false (in $\bb{R}$, for example) if the word "compact" is replaced by "closed" or by "bounded".
\end{problem}

\begin{problem}[16.]
	Regard $\bb{Q}$, the set of all rational numbers, as a metric space, with \( d(p, q) = \abs{p - q} \). Let $E$ be the set of all \( p \in \bb{Q} \) such that \( 2 < p^2 < 3 \). Show that $E$ is closed and bounded in $\bb{Q}$ but that $E$ is not compact. Is $E$ open in $\bb{Q}$?
\end{problem}

\begin{solution}
	First we will show that $E$ is bounded.\\\\
	Consider the point $q  = 1.5$ and the neighbourhood around it of radius $0.5$. We can show that $E$ is contained in that neighbourhood, and is thus bounded. We prove this by proving that any point not contained in the neighbourhood is not contained in the set as below.\\\\
	If \( d(p, q) > 0.5 \) then we have
	\[
		\abs{p - 1,5} > 0.5
	\]
	which gives us
	\( p - 1.5 > 0.5 \) or \( p - 1.5 < -0.5 \).\\\\
	Then we must have \( p > 2 \) or \( p < 1\) and thus \( p^2 < 1 \) or \( p^2 > 4 \). This gives us \( p \notin E \). Thus, we've shown that  $E$ is bounded.
	Now we will show that $E$ is closed by proving that any point that is not in $E$ cannot be a limit point of $E$.\\\\
	If \( p \notin E\) then we must have \( p^2 < 2 \) or \( p^2 > 3 \). In each of these cases e can construct a neighbourhood around $p$ that does not include any points of $E$.\\\\
	By the construction in chapter 1 to prove that the rationals do not have the least upper bound property, we know that there exists a positive rational $r$ such that for any $p$ with \( p^2 < 2 \) we have \( (p + r)^2 < 2 \). Specifically,
	\[
		r = \min{(\frac{2 - p^2}{2p + 1}, 1)}
	\]
	gives us this property. Moreover, since $r$ is positive and the square is an increasing function for positive reals, this gives us \( q^2 < 2 \) for any \( q \in (p - r, p + r) \). Thus $N_r(p)$ contains no points of $E$ making $p$ not a limit point of $E$ for any $p$ with \( p^2 < 2 \).\\\\
	We have a similar proof to show that $p$ with \( p^2 > 3 \) also cannot be a limit point of $E$. Particularly, we use a similar construction for $r$ such that the $N_r(p)$ contains no points of $E$ whenever \( p^2 > 3 \).\\\\
	Consider 
	\[
		r = \min{(\frac{p^2 - 3}{2p + 1}, 1)}
	\]
	for some $p$ with \( p^2 > 3 \).
	Then
	\[
		(p - r)^2 = p^2 - r(2p - r) > p^2 - r(2p + 1)
	\]
	since \( -r < 1 \). Substituting $r$ gives us
	\[
		(p - r)^2 > p^2 - (\frac{p^2 - 3}{2p + 1})(2p + 1) = p^2 - (p^2 - 3) = 3.
	\]
	That is
	\[
		(p - r)^2 > 3.
	\] \\
	Again, since $r$ is positive, and the square function is increasing, this gives us this gives us \( q^2 > 3 \) for any \( q \in (p - r, p + r) \). Thus $N_r(p)$ contains no points of $E$ making $p$ not a limit point of $E$ for any $p$ with \( p^2 > 3 \).\\\\
	Since we've exhausted all the cases for rationals not in $E$ (there is no rational $p$ with \( p^2 = 2 \) or \( p^2 = 3 \)) and shown all of them not to be limit points, we've shown that all the limit points of $E$ must be in $E$, and thus, $E$ must be closed.\\\\
	Finally, now we will show that $E$ is compact. Consider the cover of $E$ given by $\set{G_p}$ where \( G_p = (p - r_p, p + r_p) \) such that \( [p - r_p, p + r_p] \subseteq E \) for all \( p \in E \). Note that such an $r_p$ exists for each \( p \in E \) by a similar construction to that which we used earlier.\\\\
For any finite subset of $\set{G_p}$, say $S$, there must be some maximum $p + r_p$ chosen among all the $G_p$ in $S$. Then that maximum \( p + r_p \notin \bigcup_{G_p \in S} G_p \) but \( p + r_p \in E \) making $S$ not a finite subcover, and thus, making $E$ not compact.
\end{solution}

\begin{problem}[17.]
	Let $E$ be the set of all \( x \in [0, 1] \) whose decimal expansion contains only the digits $4$ and $7$. Is $E$ countable? Is $E$ dense in $[0, 1]$? Is $E$ compact? Is $E$ perfect?
\end{problem}

\begin{problem}[18.]
	Is there a nonempty perfect set in $\bb{R}$ which contains no rational number?
\end{problem}

\begin{problem}[19.]
	\begin{enumerate}
		\item If $A$ and $B$ are disjoint closed sets in some metric space $X$ prove that they are separated.
		\item Prove the same for disjoint open sets.
		\item Fix $p \in X$, $\delta > 0$, define $A$ to be the set of all $q \in X$ for which $d(p, q) < \delta$, define $B$ similarly, with $>$ in place of $<$. Prove that $A$ and $B$ are separated.
		\item Prove that every connected metric space with at least two points is uncountable.
	\end{enumerate}
\end{problem}

\begin{problem}[20.]
	Are closures and interiors of connected sets always connected?
\end{problem}

\begin{problem}[21.]
	Let $A$ and $B$ be separated subsets of some $\bb{R}^k$, suppose $\bvec{a} \in A$, $\bvec{b} \in B$, and define
	\[
		\bvec{p}(t) = (1 - t) \bvec{a} + t \bvec{b}
	\]
	for $t \in \bb{R}$. Put $A_0 = \bvec{p}^{\text{pre}}(A)$, $B_0 = \bvec{p}^{\text{pre}}(B)$. (Thus, $t \in A_0$ if and only if $\bvec{p}(t) \in A$).
	\begin{enumerate}
		\item Prove that $A_0$ and $B_0$ are separated subsets of $\bb{R}$.
		\item Prove that there exists $t_0 \in (0, 1)$ such that $\bvec{p}(t_0) \notin A \cup B$.
		\item Prove that every convex subset of $\bb{R}^k$ is connected.
	\end{enumerate}
\end{problem}

\begin{problem}[22.]
	A metric space is called \textit{separable} if it contains a countable dense subset. Show that $\bb{R}^k$ is separable.
\end{problem}

\begin{problem}[23.]
	A collection $\set{V_\alpha}$ of open subsets of $X$ is said to be a \textit{base} for $X$ if the following is true: for every $x \in X$ and every open set $G \subset X$ such that $x \in G$, we have $x \in V_\alpha \subset G$ for some $\alpha$. In other words, every open set in $X$ is the union of a subcollection of $\set{V_\alpha}$. \\\\
	Prove that every separable metric space has a countable base.
\end{problem}

\begin{problem}[24.]
	Let $X$ be a metric space in which every infinite subset has a limit point. Prove that $X$ is separable.
\end{problem}

\begin{problem}[25.]
	Prove that every compact metric space $K$ has a countable base, and that $K$ is therefore separable.
\end{problem}

\begin{problem}[26.]
	Let $X$ be a metric space in which every infinite subset has a limit point. Prove that $X$ is compact.
\end{problem}

\begin{problem}[27.]
	Define a point $p$ in a metric space $X$ to be a \textit{condensation point} of a set $E \subset X$ if every neighbourhood of $p$ contains uncountable many points of $E$. \\\\
	Suppose $E \subset \bb{R}^k$, $E$ is uncountable, and let $P$ be the set of all condensation points of $E$. Prove that $P$ is perfect and that at most countably many points of $E$ are not in $P$. In other words, show that $P^c \cap E$ is at most countable.
\end{problem}

\begin{problem}[28.]
	Prove that every closed set in a separable metric space is the union of a (possibly empty) perfect set and a set which is at most countable. (\textit{Corollary:} every countable closed set in $\bb{R}^k$ has isolated points.)
\end{problem}

\begin{problem}[29.]
	Prove that every open set in $\bb{R}$ is the union of an at most countable collection of disjoint segments.
\end{problem}

\begin{problem}[30.]
	Imitate the proof of theorem 2.43 to obtain the following result:\\\\
	If $\bb{R}^k = \bigcup_{n \in \bb{N}} F_n$, where each $F_n$ is a closed subset of $\bb{R}^k$, then at least one $F_n$ jas a nonempty interior.\\\\
	\textit{Equivalently} if $G_n$ is a dense open subset of $\bb{R}^k$ for $n \in \bb{N}$, then $\bigcap_{n \in \bb{N}} G_n$ is not empty. (In fact, it is dense in $\bb{R}^k$)
	
\end{problem}

\end{document}

