\documentclass[12 pt, letterpaper]{hmcpset}
\usepackage{hw-preamble}

\name{Jayden Thadani}
\class{MATH 174 --- Representation Theory}
\assignment{Homework 1}
\duedate{28th January 2025}

\begin{document}

\begin{problem}
	Write a math autobiography.
\end{problem}

\begin{solution}
	I first learnt that there was more to math than what I was learning in school when I started watching math YouTube videos online during the pandemic. I remember watching videos about how the Borsuk-Ulam theorem from algebraic topology could help divide the jewels on a necklace evenly, or how colliding blocks could compute the number $\pi$. From there, at some point, I found myself learning linear algebra and analysis, and then I just kept going.

	From \emph{Linear Algebra Done Right} to \emph{Principles of Mathematical Analysis}, from \emph{Calculus on Manifolds} to \emph{An Infinitely Large Napkin}, math books became my guides. I read vociferously, rewrote and understood every definition, result, and proof I saw, and tried hard to solve every exercise I came across. In constantly getting stuck, or coming up against things that I didn't understand, I realised that I needed a mentor. While my teachers at school weren't always able to help me with the topics I was trying to learn, I was fortunate to find others who could. 

	Though my grandfather had never been  an academic, nor did he understand the math that I wanted to do, he had always impressed upon me the importance of education. He was delighted to know that I was considering a career that would lead me to work in education. As a volunteer for a nearby of coalition of Sindhi colleges, he helped me get in touch with professors at these colleges who could guide me. Two professors --- Dr Bhandarkar and Dr Singh were kind enough to help me. They could help me get unstuck whenever I ran into problems in what I was working on, but more importantly, they showed me the elegance in other areas of math I never would have considered so interesting.

	Thanks to my grandfather, I also started considering more seriously the importance of being a mathematician as an educator. In school, I had co-founded a queer-straight alliance organisation SAFE (Students Advocate For Equality). Even more than student bullying, the most important issue that we seemed to face was that queer students felt excluded from the classroom, sometimes even explicitly. It's only then that I began to really think about the mathematician I want to be. I know now that I want to contribute to math, not just by proving theorems, but also by bringing more queer students into math. With so many niche math fields so far removed from immediate real world applications, it seems to be that the best way for mathematicians to contribute to the world is by combatting the social issues that math departments face. Thus, when I applied to colleges, I chose those which had a strong emphasis on undergraduate involvement and thriving queer communities --- colleges exactly like Mudd.

	Here at Mudd, I've been fortunate to have many mentors --- Professors Goins, Kagey, and Orrison to name just a few. In working with the Prison Education Project and organisations like Gender Equity in Math, doing research in combinatorics, and taking so many incredible math courses, I've been able to work towards my goals. This semester, I'm excited to be learning even more math, and to work with PEP on curriculum development to create more math-focused content for students.
\end{solution}

\end{document}
